---
title: "Cookbook"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cookbook}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r cook1 , include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

<!-- chunk labels need to be unique to not get conflicts when rendering Rmd KRI report -->

# Introduction

This vignette contains sample code showing how to use the [gsm](https://github.com/Gilead-BioStats/gsm) extension [gsm.simaerep](https://github.com/IMPALA-Consortium/gsm.simaerep)
using sample data from [clindata](https://github.com/Gilead-BioStats/clindata).

In order to familiarize yourself with the `gsm` package, please refer to the [gsm cookbook](https://gilead-biostats.github.io/gsm/articles/Cookbook.html).

## Installation

```{r cook2, eval=FALSE}
install.packages("pak")
pak::pak("Gilead-BioStats/clindata")
pak::pak("Gilead-BioStats/gsm.core")
pak::pak("Gilead-BioStats/gsm.mapping")
pak::pak("Gilead-BioStats/gsm.kri")
pak::pak("Gilead-BioStats/gsm.reporting")
pak::pak("IMPALA-Consortium/gsm.simaerep")
```

## Load

```{r cook3}
suppressPackageStartupMessages(library(dplyr))
library(gsm.core)
library(gsm.mapping)
library(gsm.kri)
library(gsm.reporting)
library(gsm.simaerep)
```

# {gsm.simaerep} Functions

`simaerep` expects the cumulative count of numerator events per denominator event per subject as input.

In this example we we are calculating the cumulative AE count per visit per patient per site.

```{r cook4}
dfInput <- Input_CumCount(
  dfSubjects = clindata::rawplus_dm,
  dfNumerator = clindata::rawplus_ae,
  dfDenominator = clindata::rawplus_visdt %>% dplyr::mutate(visit_dt = lubridate::ymd(visit_dt)),
  strSubjectCol = "subjid",
  strGroupCol = "siteid",
  strGroupLevel = "Site",
  strNumeratorDateCol = "aest_dt",
  strDenominatorDateCol = "visit_dt"
)

dfInput %>%
  dplyr::filter(max(Numerator) > 1, .by = "SubjectID") %>%
  head(25) %>%
  knitr::kable()
```

Now we can analyze the data using `Analyze_Simaerep()` and add flags with `Flag_Simaerep()` which adds a Score between -1 and 1. Positive values indicate the over-reporting probability and negative values indicate the under-reporting probability.

`ExpectedNumerator` is the number of expected AEs for a site with the same patient configuration.
`ScoreMult` is Score with applied multiplicity correction.

```{r cook5}
dfAnalyzed <- Analyze_Simaerep(dfInput)
dfFlagged <- Flag_Simaerep(dfAnalyzed, vThreshold = c(-0.99, -0.95, 0.95, 0.99))


dfFlagged %>%
  arrange(Score) %>%
  head(25) %>%
  knitr::kable()
```

```{r cook6}
dfFlagged %>%
  arrange(Score) %>%
  tail(5) %>%
  knitr::kable()
```

These results are compatible with the `gsm` package for visualization.

*`simaerep* scores represent are related to the metric ratio do not use a metric based threshold for flagging. Therefore we do not need to calculate boundaries to pass to the plotting function.

```{r cook7}
gsm.kri::Visualize_Scatter(
  dfFlagged,
  dfBounds = NULL,
  strGroupLabel = "GroupLevel",
  strUnit = "Visits"
)
```

```{r cook8}
Widget_ScatterPlot(
  dfFlagged,
  dfBounds = NULL,
  bDebug = FALSE
)
```

```{r cook9}
Widget_BarChart(
  dfFlagged
)
```

To compare we can also use the Score with applied multiplicity correction. For this we need to lower the threshold to be less sensitive to get a similar readout. Here we can see that overall the multiplicity correction dampens the score values and reduces the number of flagged sites. Simulation studies with {simaerep} have shown that multiplicity correction decreases detection rates. Nevertheless when monitoring a limited number of studies with many sites a sharper signal might be preferred.

```{r cook9a}
dfFlagged_Mult <- Flag_Simaerep(
  dfAnalyzed %>%
    mutate(Score = ScoreMult),
  vThreshold = c(-0.95, -0.75, 0.75, 0.95)
  )

Widget_BarChart(
  dfFlagged_Mult
)

```



# Report Building

We can create a workflow to create the `gsm` KRI report.

## Mapping

```{r cook10, message=FALSE}
lRaw <- list(
  Raw_SUBJ = clindata::rawplus_dm,
  Raw_AE = clindata::rawplus_ae,
  Raw_VISIT = clindata::rawplus_visdt,
  Raw_PD = clindata::ctms_protdev,
  Raw_ENROLL = clindata::rawplus_enroll,
  Raw_SITE = clindata::ctms_site,
  Raw_STUDY = clindata::ctms_study,
  Raw_SDRGCOMP = clindata::rawplus_sdrgcomp
)


mapping_wf <- gsm.core::MakeWorkflowList(
  strNames = NULL,
  strPath = system.file("workflow/1_mappings", package = "gsm.simaerep"),
  strPackage = NULL
)

lIngest <- gsm.mapping::Ingest(lRaw, gsm.mapping::CombineSpecs(mapping_wf))

lMapped <- gsm.core::RunWorkflows(lWorkflows = mapping_wf, lData = lIngest)
```

## Metrics

```{r cook11, message=FALSE}
metrics_wf <- gsm.core::MakeWorkflowList(
  strNames = NULL,
  strPath = system.file("workflow/2_metrics", package = "gsm.simaerep"),
  strPackage = NULL
)

lAnalyzed <- gsm.core::RunWorkflows(lWorkflows = metrics_wf, lData = lMapped)
```

## Report Generation - Workflow


```{r cook12, message=FALSE}

reporting_wf <- gsm.core::MakeWorkflowList(
  strNames = NULL,
  strPath = system.file("workflow/3_reporting", package = "gsm.simaerep"),
  strPackage = NULL
)


lReport <- gsm.core::RunWorkflows(
  lWorkflows = reporting_wf, 
  lData = c(
    lMapped,
    list(
      lAnalyzed = lAnalyzed,
      lWorkflows = metrics_wf
    )
  )
)

module_wf_gsm <- gsm.core::MakeWorkflowList(
  strNames = NULL,
  strPath = system.file("workflow/4_modules", package = "gsm.simaerep"),
  strPackage = NULL
)

# we cannot set a dynamic link to the report path in the yaml files
report_path <- system.file("report", "Report_KRI.Rmd", package = "gsm.simaerep")
n_steps <- length(module_wf_gsm$report_kri_site$steps)
module_wf_gsm$report_kri_site$steps[[n_steps]]$params$strInputPath <- report_path

lModule <- gsm.core::RunWorkflows(module_wf_gsm, lReport)
```

## Report Generation - Script

```{r cook13, warning=FALSE, message=FALSE}

dfMetrics <- gsm.reporting::MakeMetric(lWorkflows = metrics_wf)

lAnalyzed <- gsm.core::RunWorkflows(lWorkflows = metrics_wf, lData = lMapped)


dfResults <- gsm.reporting::BindResults(
  lAnalysis = lAnalyzed,
  strName = "Analysis_Summary",
  dSnapshotDate = Sys.Date(),
  strStudyID = "ABC-123"
)

dfGroups <- dplyr::bind_rows(
  lMapped$Mapped_STUDY,
  lMapped$Mapped_SITE,
  lMapped$Country
)

dfBounds <- gsm.reporting::MakeBounds(
  dfResults = dfResults,
  dfMetrics = dfMetrics
)

# we use a different tooltip for the simaerep charts
lCharts_Identity <- gsm.kri::MakeCharts(
  dfResults = dfResults %>%
    filter(GroupLevel == "Site"),
  dfMetrics = dfMetrics %>%
    filter(GroupLevel == "Site", AnalysisType %in% c("identity", "binary")),
  dfGroups = dfGroups,
  dfBounds = NULL,
  bDebug = FALSE,
  resultTooltipKeys = c(
        "ExpectedNumerator",
        "Score",
        "Metric",
        "Numerator",
        "Denominator"
  )
)

lCharts_Rate <- gsm.kri::MakeCharts(
  dfResults = dfResults %>%
    filter(GroupLevel == "Site"),
  dfMetrics = dfMetrics %>%
    filter(GroupLevel == "Site", AnalysisType == "rate"),
  dfGroups = dfGroups,
  dfBounds = dfBounds,
  bDebug = FALSE
)

lCharts <- c(
  lCharts_Identity,
  lCharts_Rate
)

gsm.kri::Report_KRI(
  lCharts = lCharts,
  dfResults = dfResults,
  dfGroups = dfGroups,
  dfMetrics = dfMetrics,
  strOutputFile = "report_kri_site.html",
  strInputPath = system.file("report", "Report_KRI.Rmd", package = "gsm.simaerep")
)
```

